---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Weyrich23

<!-- badges: start -->
<!-- badges: end -->


You will find in Weyrich23 the code used for the manuscript
vullioud et al. 2024. Please contact the corresponding author to get access to the full dataset. To
explore the script you can use the trimmed dataset available with the
package. This package is not designed for general use.

## Installation

You can install the development version of Weyrich23 from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("vullioud/Weyrich23")
```

## Analysis
The script for the analysis is available as a vignette and can be access
as follow:

```{r example,eval=FALSE}
library(Weyrich23)
browseVignettes(package = "Weyrich23")
```

The code for the functions used for the analysis is available can be
looked at here on github under the /R files.
