---
title: "analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
The scrit here provides all the code use for the analysis of the paper Weyrich et 
al. 2023. A truncated dataset (data(full_300_ws_tables_1000) is provided to try
the first part of the code. It consists of the first 1000 windows of the set, 
The code using the "path_to_annotation_file" argument is not runnable as such.
Please contact the corresponding author for the annotation file and the full 
data set. 


```{r setup, echo = FALSE, eval = FALSE}
library(Weyrich23)
library(dplyr)
```

## fit Negative-binomial model on full data set and subsets.

```{r, find DMRs,  include =TRUE, eval = FALSE}
data("full_300_ws_tables_1000") ## for access to the full dataset please contact the corresponding author. 

set.seed(123) 
describes_samples(full_300_ws_tables_1000) 


## create sets and subsets
model_1_one_fold <- prepare_balanced_k_set(k = 1, variable_to_balance = "group", tables =full_300_ws_tables)
model_1_six_fold <- prepare_balanced_k_set(k = 6, variable_to_balance = "group", tables =full_300_ws_tables)

## fit negative binomial on the windows
model_1_one_fold <- add_DMRs_to_set(model_1_one_fold, 
                                    ull_300_ws_tables, formula = "~ group + age_group", #
                                    p_cut = 0.001) ##the cut on the FDR is done later
# can change the fitting procedure...PQL is slighly more conservative but not any better in Cross-validation. 

## extract  indexes to 
index <- dplyr::bind_cols(model_1_one_fold$DMR_train) %>% 
  dplyr::filter(coef == "groupLR" & FDR <= 0.05) %>% 
  dplyr::pull(index) ## extract indexes of DMRS windows with FDR < 0.05

## fit negative binomial on the windows in the subsets
model_1_six_fold <- add_DMRs_to_set(model_1_six_fold, full_300_ws_tables, formula = "~ group + age_group", p_cut = 0.001, index = index) ## keep the result for indexes
```


## Part 1 DMR and robustness

```{r, clean DMRs,  include =TRUE, warning=FALSE,  eval = FALSE}
data("model_1_one_fold") ## results obtained with the full dataset (= analysis)
data("model_1_six_fold") ## results obtained with the full subsets (= analysis)
data("loc_table") 
path_to_annotation_table <- system.file("extdata", "crocuta_liftoff_Hhy_ASM300989v1_addPromoter.gtf", package = "Weyrich23")

DMR_group <- create_DMR_group_table(model_1_one_fold, model_1_six_fold, path_to_annotation_table, loc_table)

## descriptive stats given in the manuscript 
## Numbers given in text (results + abstract)
## get basic stats
length(unique(DMR_group$block)) ## n unique DMRs
length(unique(DMR_group$block[DMR_group$annot_null])) # n non annot
length(unique(DMR_group$block[!DMR_group$annot_null])) # m annot
length(unique(DMR_group$block[!DMR_group$annot_null & DMR_group$coherent_block])) # n cohenrent and annot
length(unique(DMR_group$block[DMR_group$coherent_block])) # n cohenrent 
length(unique(DMR_group$block[!DMR_group$coherent_block])) # n incohenrent

## get the total number of genes and their annotation 
annotated_DMR_group <- create_annot_DMRs(DMR_group)
length(unique(annotated_DMR_group$annot.gene_id)) ## total number of gene

annot_type <- get_annotation_type(annotated_DMR_group)
table(annot_type$annot.type)

fig1(DMR_group)
fig2(DMR_group)
```

## Part 2 Mitochondria and RNA

```{r, mitochondria,  include =TRUE, warning=FALSE, eval=FALSE} 
mito_table_DNA <- compute_mean_cpm_mito(full_300_ws_tables, path_to_annotation_table) ## big files you can access the computed table stored in data
mito_table_RNA <- compute_mean_cpm_mito(RNA_table)
# data(mito_table_DNA) ## computed table available with the pacakge
# data(mito_table_RNA) ## computed table available with the pacakge


mito_table_RNA %>%
  group_by(group) %>% 
  summarise(n = length(unique(ID))) ## get n = for RNA data

### get a p-value for the delta between
##get simple p-value for delta mito vs non mito and HR vs LR
coin::wilcoxsign_test(mean_count_TMM ~ as.factor(type_of_genes), data = mito_table_DNA, distribution = "exact")  
coin::wilcoxsign_test(mean_count_TMM ~ as.factor(type_of_genes), data = mito_table_RNA, distribution = "exact") ## paired wilocoxon for the mito vs non mito genes #since the samples are the same

coin::wilcox_test(mean_count_TMM ~ as.factor(group), data = mito_table_DNA, distribution = "exact") 
coin::wilcox_test(mean_count_TMM ~ as.factor(group), data = mito_table_RNA, distribution = "exact")# non-paired wilocoxon for the mito vs non mito genes since the samples are diff

### get a p-value for the delta between
fig3(mito_table_DNA, mito_table_RNA) 

```

## Part 3 RF

```{r, rf,  include =TRUE, warning=FALSE, eval = FALSE}
data(model_1_one_fold)
data(DMR_group) 
load("data-raw/full_300_ws_tables.rda")

set.seed(123)
oob_result <- estimate_oob(DMR_group, model_1_one_fold, full_300_ws_tables, 1000) ### loop on the function to have less variation due to the random nature of the analysis. // only matter for the small n. for the random draw (table in the sup)
```

## suplementary 
```{r, sup,  include =TRUE, warning=FALSE, eval = FALSE}
data(full_300_ws_tables_1000)
fig_s2b(full_300_ws_tables_1000, save = FALSE)
```
